---
title: "Exponential service time tandem queues"
author: "Tamar Loach"
date: "31 October 2016"
output: pdf_document
---

## Exponentially distributed service times

Most of the theory we've looked at for MM1 queues in series assumes exponential service times at each station. We want to simulate such queues, with a view to exploring changes in system behaviour that occur under different, application-motivated service time distrubutions. With this in mind, we look at expionential service time distributions as a comparison to theory is tractable, generating these distributions from a uniform distribution such that we can reuse this and retain the same nonlinear correlation structure when it comes to testing alternate service times distributions.


```{r}
source('Queues_Function.R')

T <- 10000
lambda <- 0.5  # param of the poisson process arrivals

# Poisson process arrivals
set.seed(13)
n <- rpois(1, lambda*T)
cat("Number of patients is: ", n)
arrival_times <- sort(c(0, runif(n - 1, min = 0, max = T)))

# Specify parameters for uniform service times
param_unif_df <- data.frame(min = c(0.51, 0.52, 0.53, 0.54),
                            max = c(0.55, 0.56, 0.57, 0.58)
                            )
# Calculate the means associated with these
param_unif_df$param <- 2/(param_unif_df$max + param_unif_df$min)

# Generate uniforms and rescale according to specified parameters
standard_unif_df <- matrix(runif(n*nrow(param_unif_df)), nrow=n)
serving_times_unif <- sapply(seq_len(ncol(standard_unif_df)), function(i){
  param_unif_df$min[i] + standard_unif_df[,i]*(param_unif_df$max[i] - param_unif_df$min[i])
})

# Inverse transform to get exp service times
serving_times_exp <- sapply(seq_len(ncol(standard_unif_df)), function(i){
  - log(1-standard_unif_df[,i])/param_unif_df$param[i]
  })

# Generate queues with the specified arrival and service dists
queues_unif_list <- QueuesFunction(serving_times_unif, arrival_times)
queues_exp_list <- QueuesFunction(serving_times_exp, arrival_times)

```

```{r}
burn_in_time = T/10
queues_unif = queues_unif_list$queues_list
queues_exp = queues_exp_list$queues_list

queues_unif = lapply(queues_unif, function(x) x[x$time <=T & x$time > burn_in_time, ])
queues_exp = lapply(queues_exp, function(x) x[x$time <=T & x$time > burn_in_time, ])

```



Plot the number of people in each queue over time


```{r}
source('Queues_Plots.R')

PlotNrPeople(queues_unif)
PlotNrPeople(queues_exp)

```


Plot the hist of queue legnths for each queue comparing to geometric dist

```{r}
source('Queues_Plots.R')

rho = lambda/param_unif_df$param

plot = HistNrPeopleWithGeom(queues_unif, 1 - rho)
plot

plot = HistNrPeopleWithGeom(queues_exp, 1 - rho)
plot


```
